Appendix 1: Quantifying Drought

### 2021-03-15 analysis/paper/vital rate models.Rmd
add sentence wrap option to YAML

---
title: "FLM for growth as function of SPEI"
author: "Eric R. Scott"
date: "Date Initiated"
output: 
  html_notebook: 
    highlight: kate
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: yes
---

@@

- ~~Switch measure of plant size to shoots * height~~
- ~~Change lag to 36 months (3 years)~~
- Produce ridge plot
- Heatmap for growth should be ∆size
- try select = TRUE (after you understand it better? blog about it first)

@@

# Survival
`surv` is 1 in `year`s when a plant is alive and 0 in the first `year` when the plant is dead.  These GAMs model the survival of a plant (over the past year) as a function of size in the previous census and SPEI values starting at february of the current year going back 36 months.
## Fit GAMs

@@

```{r}
s_post_1ha <-
  bam(surv ~ 
        s(log_size_prev, bs = "cr") +
        s(plot, bs = "re") +
        # s(plot, log_size_prev, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_1ha,
      method = "REML",
      select = TRUE,
      cluster = cl)
```
```{r}
#histogram
plot(density(qresid(s_post_1ha)))
#QQ plot
qqnorm(qresid(s_post_1ha)); qqline(qresid(s_post_1ha))
#Fitted vs. residuals
scatter.smooth(predict(s_post_1ha, type = "response"), qresid(s_post_1ha), col = "grey")
```

@@

### Setup for all crossbasis plots
Ok, I want to add a lil strip on the bottom that shows dry and wet seasons.  Wet = November-may, dry = june-october.

@@

### Get plot data
`dist` isn't working yet in `smooth_estimates`, but that's ok because I don't need CI for these plots (yet).

@@

### Difference
Try plotting a raster of the difference between CF and 1-ha

@@

# Growth
For mature plants that are alive...

@@

AIC for... 
Gaussian: ~40909.27. looks leptokurtic
scaled t: ~37284.47.  Waaay better

@@

ggplot(g_1ha_eval1, aes(x = log_size_prev)) +
  geom_line(aes(y = est)) +
  # geom_abline(slope = 1, linetype = 2, color = "red") +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), alpha = 0.25) +
  geom_rug(data = model.frame(g_post_1ha)) +
  theme_bw() +
  labs(x = TeX("$log(size_t)$"), y = TeX("$log(size_{t+1})"), title = "1-ha fragments")
ggsave(here("analysis", "figures", "growth_1ha_size.png"))
```
```{r}
g_cf_eval1 <-
  smooth_estimates(g_post_cf, smooth = "log_size_prev", unconditional = TRUE) %>% 
  add_confint() %>% 
  mutate(across(c(est, lower_ci, upper_ci), ~.x + coef(s_post_cf)[1]))

@@

### Combine
combine
```{r}
(g_cf_rotated + theme(axis.title.x = element_blank())) / 
  (g_1ha_rotated) + 
  plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = "a", tag_suffix = ")") & 
  labs(title = NULL) #remove plot titles
ggsave(here("analysis", "figures", "growth_both_spei.png"), width = 7, height = 5)
```
### Difference

@@

```{r}
plogis(coef(f_post_1ha)[1])
plogis(coef(f_post_cf)[1])
```
Overall probability of flowering higher in continuous forest (0.008) compared to 1ha fragments (0.003)

@@

```{r}
bind_rows("1-ha" = f_1ha_eval1, "CF" = f_cf_eval1, .id = "habitat") %>% 
  ggplot(aes(x = log_size_prev)) +
  geom_line(aes(y = est, color = habitat)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci, fill = habitat), alpha = 0.25) +
  # stat_summary_bin(geom = "point", fun = mean, aes(y = flwr), data = model.frame(f_post_cf)) +
  geom_rug(data = bind_rows(
    "1-ha" = model.frame(f_post_1ha),
    "CF" = model.frame(f_post_cf),
    .id = "habitat")) +
  theme_bw() +
  labs(x = TeX("$log(size_t)$"), y = "P(flowering)")
ggsave(here("analysis", "figures", "flwr_both_size.png"))
```
```{r paged.print=FALSE}
demog_full %>% filter(ht < 1)
demog_full %>% 
  # filter(log_size > 7) %>%
  filter(shts >= 4) %>% 
  group_by(habitat) %>% 
  summarise(flower_prob = mean(flwr, na.rm = TRUE))

@@

## Crossbasis plot
### Get plot data
`dist` isn't working yet in `smooth_estimates`, but that's ok because I don't need CI for these plots (yet).

@@

### CF
```{r}
f_cf_rotated <-
  ggplot(f_cf_eval, aes(y = spei_history, x = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.001) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c("P(flowering)", option = "viridis", limits = flwr_lims) +
  scale_x_continuous(
    "lag (months before census)",
    breaks = breaks,
    expand = c(0, 0)
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0)))+ #leave room for season bar at bottom
  labs(title = "Continuous forest") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL
```
### Combine
combine
`
```{r}
  (f_cf_rotated + theme(axis.title.x = element_blank())) / 
  (f_1ha_rotated) + 
  plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = "a", tag_suffix = ")") & 
  labs(title = NULL) #remove plot titles
ggsave(here("analysis", "figures", "flwr_both_spei.png"), width = 7, height = 5)
```
### Difference

>>>>

---
title: "FLM for growth as function of SPEI"
author: "Eric R. Scott"
date: "Date Initiated"
output: 
  html_notebook: 
    highlight: kate
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: yes
editor_options:
  markdown:
    wrap: sentence
    canonical: true
---

@@

-   ~~Switch measure of plant size to shoots \* height~~
-   ~~Change lag to 36 months (3 years)~~
-   Produce ridge plot
-   Heatmap for growth should be ∆size
-   try select = TRUE (after you understand it better? blog about it first)

@@

# Survival

`surv` is 1 in `year`s when a plant is alive and 0 in the first `year` when the plant is dead.
These GAMs model the survival of a plant (over the past year) as a function of size in the previous census and SPEI values starting at february of the current year going back 36 months.
\#\# Fit GAMs

@@

```{r}
s_post_1ha <-
  bam(surv ~ 
        s(log_size_prev, bs = "cr") +
        s(plot, bs = "re") +
        # s(plot, log_size_prev, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_1ha,
      method = "REML",
      select = TRUE,
      cluster = cl)
```

```{r}
#histogram
plot(density(qresid(s_post_1ha)))
#QQ plot
qqnorm(qresid(s_post_1ha)); qqline(qresid(s_post_1ha))
#Fitted vs. residuals
scatter.smooth(predict(s_post_1ha, type = "response"), qresid(s_post_1ha), col = "grey")
```

@@

### Setup for all crossbasis plots

Ok, I want to add a lil strip on the bottom that shows dry and wet seasons.
Wet = November-may, dry = june-october.

@@

### Get plot data

`dist` isn't working yet in `smooth_estimates`, but that's ok because I don't need CI for these plots (yet).

@@

### Difference

Try plotting a raster of the difference between CF and 1-ha

@@

# Growth

For mature plants that are alive...

@@

AIC for... Gaussian: \~40909.27.
looks leptokurtic scaled t: \~37284.47.
Waaay better

@@

ggplot(g_1ha_eval1, aes(x = log_size_prev)) +
  geom_line(aes(y = est)) +
  # geom_abline(slope = 1, linetype = 2, color = "red") +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), alpha = 0.25) +
  geom_rug(data = model.frame(g_post_1ha)) +
  theme_bw() +
  labs(x = TeX("$log(size_t)$"), y = TeX("$log(size_{t+1})"), title = "1-ha fragments")
ggsave(here("analysis", "figures", "growth_1ha_size.png"))
```

```{r}
g_cf_eval1 <-
  smooth_estimates(g_post_cf, smooth = "log_size_prev", unconditional = TRUE) %>% 
  add_confint() %>% 
  mutate(across(c(est, lower_ci, upper_ci), ~.x + coef(s_post_cf)[1]))

@@

combine

```{r}
(g_cf_rotated + theme(axis.title.x = element_blank())) / 
  (g_1ha_rotated) + 
  plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = "a", tag_suffix = ")") & 
  labs(title = NULL) #remove plot titles
ggsave(here("analysis", "figures", "growth_both_spei.png"), width = 7, height = 5)
```

### Difference

@@

```{r}
plogis(coef(f_post_1ha)[1])
plogis(coef(f_post_cf)[1])
```

Overall probability of flowering higher in continuous forest (0.008) compared to 1ha fragments (0.003)

@@

```{r}
bind_rows("1-ha" = f_1ha_eval1, "CF" = f_cf_eval1, .id = "habitat") %>% 
  ggplot(aes(x = log_size_prev)) +
  geom_line(aes(y = est, color = habitat)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci, fill = habitat), alpha = 0.25) +
  # stat_summary_bin(geom = "point", fun = mean, aes(y = flwr), data = model.frame(f_post_cf)) +
  geom_rug(data = bind_rows(
    "1-ha" = model.frame(f_post_1ha),
    "CF" = model.frame(f_post_cf),
    .id = "habitat")) +
  theme_bw() +
  labs(x = TeX("$log(size_t)$"), y = "P(flowering)")
ggsave(here("analysis", "figures", "flwr_both_size.png"))
```

```{r paged.print=FALSE}
demog_full %>% filter(ht < 1)
demog_full %>% 
  # filter(log_size > 7) %>%
  filter(shts >= 4) %>% 
  group_by(habitat) %>% 
  summarise(flower_prob = mean(flwr, na.rm = TRUE))

@@

### Get plot data

`dist` isn't working yet in `smooth_estimates`, but that's ok because I don't need CI for these plots (yet).

@@

```{r}
f_cf_rotated <-
  ggplot(f_cf_eval, aes(y = spei_history, x = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.001) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c("P(flowering)", option = "viridis", limits = flwr_lims) +
  scale_x_continuous(
    "lag (months before census)",
    breaks = breaks,
    expand = c(0, 0)
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0)))+ #leave room for season bar at bottom
  labs(title = "Continuous forest") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL
```

combine \`

```{r}
  (f_cf_rotated + theme(axis.title.x = element_blank())) / 
  (f_1ha_rotated) + 
  plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = "a", tag_suffix = ")") & 
  labs(title = NULL) #remove plot titles
ggsave(here("analysis", "figures", "flwr_both_spei.png"), width = 7, height = 5)
```

### Difference

### 2021-03-15 analysis/supplementary-materials/appendix_SPEI.Rmd
add sentence wrap option to YAML

---
title: "Appendix 1: Quantifying Drought"
author: "Eric R. Scott"
date: '2020-09-13'
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/ecology.csl" # Insert path for the bib-style
---

>>>>

---
title: "Appendix 1: Quantifying Drought"
author: "Eric R. Scott"
date: '2020-09-13'
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/ecology.csl" # Insert path for the bib-style
editor_options:
  markdown:
    wrap: sentence
    canonical: true
---
